### 2014-09-11 paper/paper.tex
started intro

\documentclass[useAMS,usenatbib]{mn2e}
\usepackage{epsfig,rotate,graphicx}
\usepackage[fleqn]{amsmath}
\usepackage{subfigure}
\usepackage{lscape}
\usepackage{bm}

\newcommand{\p}{\partial}
\newcommand{\mnras}{MNRAS}
\newcommand{\apj}{ApJ}
\newcommand{\aap}{A\&A}
\newcommand{\apjl}{ApJL}
\newcommand{\araa}{ARAA}
\newcommand{\dd}{\delta}
\newcommand{\adot}{\dot{a}}
\newcommand{\be}{\begin{equation}}
\newcommand{\ee}{\end{equation}}
\newcommand{\gtrsim}{\;\raisebox{-.8ex}{$\buildrel{\textstyle>}\over\sim$}\;}
\newcommand{\lesssim}{\; \raisebox{-.8ex}{$\buildrel{\textstyle<}\over\sim$}\;}
\newcommand{\anrev}{{\it ARA\&A, }}
\newcommand{\apjs}{{\it ApJS, }}
\newcommand{\icar}{{\it Icarus, }}
\newcommand{\mnr}{{\it MNRAS, }}
\newcommand{\nat}{{\it Nature, }}
\newcommand{\sci}{{\it Sci, }}
\newcommand{\ana}{{\it A\&A, }}
\newcommand{\anas}{{\it A\&AS, }}
\newcommand{\aaps}{{\it A\&AS, }}
\newcommand{\anar}{{\it A\&AR, }}
\newcommand{\prd}{{\it Phys. Rev. D }}
\newcommand{\qjl}{{\it QJRAS, }}
\newcommand{\sbar}{\bar{\sigma}}
\newcommand{\avg}[1]{\langle #1 \rangle_\phi}
\newcommand{\bu}{\bm{u}}
\newcommand{\lmax}{l_\mathrm{max}}
\newcommand{\mmax}{m_\mathrm{max}}
\newcommand{\trel}{t_\mathrm{rel}}
\newcommand{\zeus}{{\tt ZEUS-MP }}
\newcommand{\ii}{\mathrm{i}}
\newcommand{\bv}{\bm{v}}
\newcommand{\rin}{r_\mathrm{in}}
\newcommand{\rout}{r_\mathrm{out}}
\newcommand{\ciso}{c_\mathrm{iso}}
\newcommand{\tbeta}{\tilde{\beta}}
\newcommand{\teta}{\tilde{\eta}}
\newcommand{\tcool}{t_\mathrm{cool}}
\DeclareMathOperator{\erf}{erf}

\usepackage{array,booktabs,tabularx}
\newcolumntype{R}{>{\centering\arraybackslash}X} % right justified tabularx columns

\title[Gaps in non-isothermal discs]{Gap formation and stability in 
  non-isothermal protoplanetary discs}

\author[Les and Lin]{Robert Les
  $^1$\thanks{robert.les@mail.utoronto.ca} and Min-Kai Lin $^1$
  \thanks{ mklin924@cita.utoronto.ca} \\ 
$^1$Canadian Institute for Theoretical Astrophysics,  
60 St. George Street, Toronto, ON, M5S 3H8, Canada \\
}

\begin{document}

\maketitle
\begin{abstract}
%Hydrodynamic simulations of self-gravitating disc-planet interactions are presented. 
Several observations of transition discs show lopsided
dust-distributions. This is suggestive of a dust-trapping mechanism at work. A
popular explanation is the formation of a large-scale vortex at the
edge of a gap opened by a giant planet. While theoretical models
support this picture, numerical simluations have thus far
employed locally isothermal discs. On the other hand, the theory of the
vortex-forming instability was originally developed for adiabatic
discs. We generalise the study of planetary gap stability to non-isothermal
discs using customised numerical simulations of disc-planet
systems. We include in the energy equation a cooling term with cooling
timescale $t_\mathrm{cool}=\beta\Omega_k^{-1}$, where $\Omega_k$ is
the Keplerian frequency, and examine the effect of $\beta$ on the
stability of gap edges. For measurments reminsent of linear stability, gap edges stabilize with increasing $\beta$ as dominate modes and growth rates decrease in value. In the non-linear limit of planet-vortex interactions, large scale overdense vortices were found to last for 1000's of orbits in a quasi-steady state characterized by sudden disspation on the order of 100's or planet orbits. Vortex lifetimes were found to have a non-monoticly dependence on $\beta$ and an optimum value for cooling rate was found to maximize the lifetime.

%The simulations are customized to examine the stability of gaps opened by massive planets as a function of $\beta$.

\end{abstract}

\begin{keywords}
planetary systems: formation --- planetary systems:
protoplanetary discs
\end{keywords}

\section{Introduction}\label{intro}
{
\bf 
observational motivation: asymmetric dust distributions. 
theoretical motivations: generalize gap stability to include energy equation. 
}

Disk-planet interaction plays an important role in the theory of planet formation and has been invoked to explain 
sub-structures obsererved in transition disks. It is well-known that a sufficiently massive planet opens a disk gap, and
such structures have been detected. A more recent theoretical result is that planetary gaps are dynamically unstable under
certain conditions. The so-called Rossby wave instability operates in radially structured disks, and leads 
to vortex-formation.

In this work, we extend the study of planetary gap stability against vortex formation to non-isothermal discs. 
We include a fluid energy equation in the hydrodynamical equations describing disk-planet systems. We consider a simple
cooling function and examine the effect cooling rate on the stability of planetary gaps.

This paper is organised as follows.

\input{model}
\input{results}
%\input{discussion}
%\input{summary}

%\section*{Acknowledgments}

\bibliographystyle{mn2e}
\bibliography{ref}

\appendix
%\input{appendix}
%\input{appendix2}

\end{document}