### 2014-09-19 paper/model.tex
edited linear discussion

\section{Disc-planet models}\label{model}
The system is a two-dimensional non-self-gravitating gas disc orbiting
a central star of mass $M_*$.  
%self-gravitating gas disc of mass
%$M_d$ rotating about a central star of mass $M_*$. 
Cylindrical
co-ordinates $(r,\phi,z)$ centred on the star is adopted. The frame is   
non-rotating. Computational units are such that 
$G=M_*=\mathcal{R}=\mu=1$ where $G$ is the gravitational constant,
$\mathcal{R}$ is the gas constant and $\mu$ is the mean molecular
weight. %% \cite{lin12}

The disc is governed by the standard fluid equations  
\begin{align}\label{3d_gov_eq}
  &\frac{\p\Sigma}{\p t}+\nabla\cdot(\Sigma \bv)=0, \\
  & \frac{\p\bv}{\p t}+\bv\cdot\nabla\bv= -\frac{1}{\rho}\nabla p 
  - \nabla{\Phi} + \bm{f}_\nu,\\
  & \frac{\p e}{\p t} + \nabla\cdot(e\bv) = -p\nabla\cdot\bv +
  \mathcal{H} - \mathcal{C}, %\\
%  &\nabla^2\Phi_d = 4\pi G \Sigma \delta(z),\label{poisson}
\end{align}
where $\Sigma$ is the surface density, $\bv = (v_r,v_\phi)$ the fluid
velocity, $p$ is the vertically-integrated pressure, $e=p/(\gamma-1)$ is the energy
per unit area and the adiabatic index $\gamma=1.4$ is assumed
constant.

The total potential $\Phi$ includes the stellar potential, planet potential
(described below) %, and the disc potential $\Phi_d$ which satisfies the
%Poisson equation \ref{poisson}. 
and indirect potentials to account for the non-inertial reference
frame.%  \citep[see][for explicit expressions]{masset02}
In the momentum 
equations, $\bm{f}_\nu$ represent viscous forces whose magnitude is 
characterised by a constant kinematic viscosity parameter
$\nu$. However, we will be considering effectively inviscid discs by
adopting small values of $\nu$ .

\subsection{Heating and cooling}
In the energy equation, the heating term $\mathcal{H}$ is defined as 
\begin{align}
  \mathcal{H} \equiv Q^+ - Q^+_i\frac{\Sigma}{\Sigma_i}, 
\end{align}
where $Q^+$ represents viscous heating and subscript $i$ denotes
evaluation at $t=0$. The cooling term $\mathcal{C}$ is defined as
\begin{align}
  \mathcal{C} \equiv \frac{1}{t_c}\left(e -
  e_i\frac{\Sigma}{\Sigma_i}\right),  
\end{align}
where $t_c = \beta\Omega_k^{-1}$ is the cooling time,
$\Omega_k=\sqrt{GM/r^3}$ is the Keplerian frequency and $\beta$ is an
input parameter. This cooling prescription allows one 
to explore the full range of thermodynamic response of the disc in a 
systematic way: $\beta\ll1$ is a locally isothermal disc while
$\beta\gg1$ is an adiabatic disk.

Note that the energy source terms
have been chosen to be absent at $t=0$, allowing the disc to be
initialised close to steady state. The $\mathcal{C}$ function attempts to restore the initial energy density (and
therefore temperature) profile. In practice, this is a cooling term at
the gap edge because disc-planet interaction leads to heating.

\subsection{Disc model and initial condition}
The disc occupies $r\in[r_\mathrm{in}, r_\mathrm{out}]$ with
$r_\mathrm{out}=25\mathrm{in}$ {\bf check: if all simulations use the
  same domain, state it here} and $\phi\in[0,2\pi]$. The initial disc is axisymmetric with
surface density profile

\begin{align}\label{initial_density}
   \Sigma(r) = \Sigma_\mathrm{ref}\left(\frac{r}{r_\mathrm{in}}\right)^{-s}
    \left[1 - \sqrt{\frac{r_\mathrm{in}}{r + H_i(\rin)}}\,\right] 
\end{align}
where the power-law index $s=2$, $H(r) = \ciso\Omega_k $ defines the disc scale-height 
where $\ciso=\sqrt{p/\Sigma}$ isothermal sound-speed. The disc aspect-ratio is defined as $h\equiv H/r$ and initially
$h=0.05$. For a non-self-gravitating disc, the surface density scale
$\Sigma_\mathrm{ref}$ is arbitrary.

%The surface density scale $\Sigma_0$ is set by  
%\begin{align}
%\Sigma_0 = \frac{h_i\rout\Omega_k^2(\rout)}{\pi G Q_o},
%\end{align}
%where $Q_o$ is the thin-disc Keplerian Toomre parameter at the outer boundary initially.

The initial azimuthal velocity $v_{\phi i}$ is set by centrifugal balance with
pressure, stellar gravity and disc gravity. The initial radial velocity is $v_{r}=3\nu/r$ 
with $\nu = \hat{\nu}\rin^2\Omega_k(\rin)$. For parameter values employed in this work, $v_{\phi}\simeq r\Omega_k$ 
and $\hat{\nu}= 10^{-9}$, so that $|v_{r}/v_{\phi}|\ll1$ and the initial 
flow is effectively only in the azimuthal direction.  With this choice
of physical viscosity $\hat{\nu}$ the only heating source is through
shock-heating (via artificial viscosity) and the $\mathcal{C}$
function when $e\Sigma<e_i\Sigma_i$.

\subsection{Planet potential}\label{planet_config}
The planet potential is given by 
\begin{align}
  \Phi_p = -\frac{GM_p}{\sqrt{|\bm{r} - \bm{r}_p|^2 + \epsilon_p^2}},
\end{align}
where $M_p$ is the planet mass and we fix $q\equiv M_p/M_*=10^{-3}$
throughout this work. This corresponds to a Jupiter-mass planet if $M_*=M_{\sun}$. 
The planet's position in the disc 
$\bm{r}_p=(r_p,\phi_p)$  and $\epsilon_p=0.5r_h$ is a softening
length with $r_h=(q/3)^{1/3}r_p$ being the Hill radius.  
%Hill radius chosen so eps doesn't change with time (scale-height does)
The planet is held on a fixed circular orbit with $ r_p = 10\rin$. This also
defines the time unit $P_0\equiv 2\pi/\Omega_k(r_p)$ used to describe results. 
%In this study only one planet mass is considered, with $q=10^{-3}$. 
%The initial orbital radius is

\section{Numerical experiments}\label{method}
The disc-planet system is evolved using the 
\texttt{FARGO-ADSG} code \citep{baruteau08, baruteau08b}. This is a modified version 
of the original \texttt{FARGO} code \citep{masset00a} to include the energy 
equation. The code employs a finite-difference scheme similar 
to the \texttt{ZEUS} code \citep{stone92}, but with a modified azimuthal transport 
algorithm to circumvent the time-step restriction set by the inner disc boundary. 
The disc is divided into $(N_r,N_\phi)$ zones in the radial and azimuthal directions, 
respectively. The grid spacing is logarithmic in radius and uniform in azimuth.

\subsection{Parametrized cooling}
In this work we only vary one control parameter: the cooling
time. %The disc parameters are $Q_o$, which characterizes the strength of self-gravity; 
%and $\beta$ which sets the cooling time. Two levels of self-gravity is considered: 
%a light disc with $Q_o=8$ and a heavy disc with $Q_o=1.5$. 
The cooling parameter $\beta$ is chosen indirectly  through the parameter
$\tilde{\beta}$ such that

\begin{align}
  t_c(r_p+x_s) = \beta\Omega_k^{-1}(r_p+x_s) = \tilde{\beta} t_{\mathrm{lib}}(r_p+x_s), 
\end{align}
where $x_s=2r_h$ is the approximate distance from the planet to its
gap edge, and $t_\mathrm{lib}$ is the time interval between successive
encounters of a fluid element at the gap edge and the planet's
azimuth. That is, we measure the cooling time in units of the time
taken for a gap edge fluid element to encounter planet-induced
shocks.

Assuming Keplerian orbital frequencies and $x_s\ll r_p$
gives $t_\mathrm{lib}\simeq 4\pi r_p/(3\Omega_{kp} x_s)$, where
$\Omega_{kp} = \Omega_k(r_p)$. Therefore   
\begin{align}\label{betatilde}
  \beta = \tilde{\beta} \frac{4\pi r_p}{3x_s} \left(1  - \frac{3x_s}{2r_p}\right), 
\end{align}
where $x_s\ll r_p$ was used again. We use $x_s = 2r_h$ in
Eq. \ref{betatilde}. For a planet mass with $q=10^{-3}$,
Eq. \ref{betatilde} then gives $\beta \simeq 23.9\tilde{\beta}$.

%giant planet gap is 2.5rh
%2rh is CR of G.E.I. 
%\subsection{Simulation procedure}
%The disc is evolved without a planet for $t\in[0,20]P_0$. The planet potential is smoothly switched on
%between $t\in[20,30]P_0$.
%%  In a separate set of calculations the planet
%% potential is turned off after $t>t_\mathrm{off}$, immediately  
%% afterwhich the surface density, energy density and velocity fields are
%% azimuthally averaged and then subject to random surface density  
%% perturbations of magnitude $10^{-2}$.

\subsection{Diagonistic measures}

\subsubsection{Generalized potential vorticity}

The generalized potential vorticity is defined as
\begin{align}
  \tilde{\eta} = \frac{\kappa^2}{2\Omega\Sigma}\times S^{-2/\gamma}, 
\end{align}
where $\kappa^2 = r^{-3}\partial_r(r^4\Omega^2)$ is the square of the
epicycle frequency, $\Omega=v_\phi/r$ is the angular speed, and
$S\equiv p/\Sigma^\gamma$ is the entropy. The first factor is the
usual potential vorticity (PV, or vortensity).

The generalized PV appears in the description of the linear stability
of radially-structured adiabatic discs \citep{lovelace99,li00}, where
the authors show an extremum in $\teta$ may lead to a dynamical
instability. In a barotropic disc, the entropy factor is absent
and the important quantity is the PV.

\subsubsection{Fourier modes} 
The Rossby wave instability is characterized by exponetially
growing pertibations. Though in this paper we do not formally consider
linear instability, modal analysis will be useful to
analyse the growth of different wavenumbers, which is associated with
the number of vortices initially formed by the RWI.

%characterize 
%azimuthal vortex numbers and overdensity values.

The Fourier transform of the time-dependent surface density is
\begin{align}\label{fouriertransform}
  \Sigma_m(r,t) = \int_{0}^{2\pi}
  \Sigma(r,\phi,t) \, \mathrm{e}^{-\mathrm{i}m\phi} \, \mathrm{d}\phi 
\end{align} 
where $m$ is the aziumuthal wave number. We define the growth rate
$\gamma$ of the $m^\mathrm{th}$ component of the surface density
through 
% and $\sigma_m$ is the complex frequency
%of the $m \mathrm{th}$ mode. Since $\sigma_m$ is a value dependent on
%$r$, a more useful diagnostic is the mode growth rate defined as  
\begin{align}\label{growth}
  \frac{d \langle|\Sigma_m|\rangle_r }{dt}= \gamma \langle|\Sigma_m|\rangle_r 
\end{align}
where %$\gamma$ is the real part of the $\sigma_m$ and 
$\langle|\cdot|\rangle_r$ denotes the average of the absolute value
over a raidal region of interest. By using Eq.~\ref{growth} the growth
rates of the unstable modes can be found from successive spatial
Fourier transforms over an appropriate period of time.

\subsubsection{Rossby number}
The Rossby number
\begin{align}
  Ro = \frac{{\bf \hat{z}} \cdot \nabla \times \bv - \langle
    {\bf\hat{z}} \cdot \nabla \times \bv
    \rangle_{\phi}}{2\langle\Omega\rangle_{\phi}},  
\end{align}
is a dimensionless measure of relative vorticity. 
% the ratio of interial to Corriolis forces. 
Here $\langle \cdot \rangle_{\phi}$ denotes an azimuthal
average. Values of $Ro<0$ correspond to anti-cyclonic rotation with
respect to the background shear and thus can be used to identify
vortices and quantify its intensity.

%{\bf {MKL.} describe vortensity and generalized vortensity, and its
%  relevance to stability. fourier mode amplitude definition and how
%  growth rates are obtained. rossby number. density
%  perturbations. (see lin 2014)
%}

%% According to previous 2D studies, the key quantitiy in describing the
%% stability of radially structured non-barotropic discs is the
%% \emph{generalized vortensity} (GV): 
%% \begin{align}
%%   \tilde{\eta}\equiv \left(\frac{\kappa^2}{2\Omega\Sigma}\right)S^{-2/\gamma}, 
%% \end{align}
%% where $S \equiv p/\Sigma^\gamma$ is defined to be the entropy. The
%% quantity in brackets, $\eta\equiv \kappa^2/2\Omega\Sigma$, is the
%% usual vortensity or potential vorticity (PV).