### 2014-10-15 paper/paper.tex
intro added, abstract edit, ref update

\documentclass[useAMS,usenatbib]{mn2e}
\usepackage{epsfig,rotate,graphicx}
\usepackage[fleqn]{amsmath}
\usepackage{subfigure}
\usepackage{lscape}
\usepackage{bm}

\newcommand{\p}{\partial}
\newcommand{\mnras}{MNRAS}
\newcommand{\apj}{ApJ}
\newcommand{\aap}{A\&A}
\newcommand{\apjl}{ApJL}
\newcommand{\araa}{ARAA}
\newcommand{\dd}{\delta}
\newcommand{\adot}{\dot{a}}
\newcommand{\be}{\begin{equation}}
\newcommand{\ee}{\end{equation}}
\newcommand{\gtrsim}{\;\raisebox{-.8ex}{$\buildrel{\textstyle>}\over\sim$}\;}
\newcommand{\lesssim}{\; \raisebox{-.8ex}{$\buildrel{\textstyle<}\over\sim$}\;}
\newcommand{\anrev}{{\it ARA\&A, }}
\newcommand{\apjs}{{\it ApJS, }}
\newcommand{\icar}{{\it Icarus, }}
\newcommand{\mnr}{{\it MNRAS, }}
\newcommand{\nat}{{\it Nature, }}
\newcommand{\sci}{{\it Sci, }}
\newcommand{\ana}{{\it A\&A, }}
\newcommand{\anas}{{\it A\&AS, }}
\newcommand{\aaps}{{\it A\&AS, }}
\newcommand{\anar}{{\it A\&AR, }}
\newcommand{\prd}{{\it Phys. Rev. D }}
\newcommand{\qjl}{{\it QJRAS, }}
\newcommand{\sbar}{\bar{\sigma}}
\newcommand{\avg}[1]{\langle #1 \rangle_\phi}
\newcommand{\bu}{\bm{u}}
\newcommand{\lmax}{l_\mathrm{max}}
\newcommand{\mmax}{m_\mathrm{max}}
\newcommand{\trel}{t_\mathrm{rel}}
\newcommand{\zeus}{{\tt ZEUS-MP }}
\newcommand{\ii}{\mathrm{i}}
\newcommand{\bv}{\bm{v}}
\newcommand{\rin}{r_\mathrm{in}}
\newcommand{\rout}{r_\mathrm{out}}
\newcommand{\ciso}{c_\mathrm{iso}}
\newcommand{\tbeta}{\tilde{\beta}}
\newcommand{\teta}{\tilde{\eta}}
\newcommand{\tcool}{t_\mathrm{cool}}
\DeclareMathOperator{\erf}{erf}

\usepackage{array,booktabs,tabularx}
\newcolumntype{R}{>{\centering\arraybackslash}X} % right justified tabularx columns

\title[Gaps in non-isothermal discs]{Gap formation and stability in 
  non-isothermal protoplanetary discs}

\author[Les and Lin]{Robert Les
  $^1$\thanks{robert.les@mail.utoronto.ca} and Min-Kai Lin $^{1,2}$
  \thanks{ minkailin@email.arizona.edu} \\ 
$^1$Canadian Institute for Theoretical Astrophysics,  
60 St. George Street, Toronto, ON, M5S 3H8, Canada \\
$^2$Department of Astronomy and Steward Observatory, University of
Arizona, 933 North Cherry Avenue, Tucson, AZ 85721, USA 
}

\begin{document}

\maketitle
\begin{abstract}
%Hydrodynamic simulations of self-gravitating disc-planet interactions are presented. 
Several observations of transition discs show lopsided
dust-distributions. A potential explanation is the formation of a
large-scale vortex acting as a dust-trap at the edge of a gap opened
by a giant planet. While numerical models of gap-edge vortices have
thus far employed locally isothermal discs, the 
theory of this vortex-forming or `Rossby wave' instability was
originally developed for adiabatic discs.  
We generalise the study of planetary gap stability to non-isothermal
discs using customised numerical simulations of disc-planet
systems where the planet opens unstable gaps. 
We include in the energy equation a simple cooling term with
cooling timescale $t_\mathrm{cool}=\beta\Omega_k^{-1}$, where $\Omega_k$ is
the Keplerian frequency, and examine the effect of $\beta$ on the
stability of gap edges and vortex lifetimes. We find increasing
$\beta$ lowers the growth rate of non-axisymmetric perturbations, and the
dominant azimuthal wavenumber $m$ decreases.  
%gap edges stabilize with increasing $\beta$ as dominate modes and
%growth rates decrease in value. 
We find a quasi-steady state consisting of one  
large-scale overdense vortex circulating the outer gap edge, typically  
lasting $O(10^3)$ orbits.  
%characterized by
%sudden disspation on the order of 100's or planet orbits.
Vortex lifetimes were found to increase with cooling times up to an
optimal value, after which vortex lifetimes decrease. This
non-monotonic dependence is qualitatively consistent with 
strictly isothermal simulations that differ in disc aspect ratios
presented in Fu et al. (2014, ApJL, 788, L41).

%a non-monoticly dependence on $\beta$ and
%an optimum value for cooling rate was found to maximize the lifetime.

%The simulations are customized to examine the stability of gaps
%opened by massive planets as a function of $\beta$.

\end{abstract}

\begin{keywords}
planetary systems: formation --- planetary systems:
protoplanetary discs
\end{keywords}

\section{Introduction}\label{intro}
%{
%\bf 
%observational motivation: asymmetric dust distributions. 
%theoretical motivations: generalize gap stability to include energy equation. 
%}
The interaction between planets and protoplantary discs plays an
important role in the theory of planet formation and disk 
evolution. Disk-planet interaction may lead to the orbital migration
of protoplanets and modify the structure of
protoplanetary disks  \citep[see][for a recent review]{baruteau13}.  
%\citep[][see \cite{baruteau13} for a recent
%  review]{goldreich79, lin79, goldreich80,lin86}.

A sufficiently massive planet can open a gap in a 
gaseous protoplanetary disc \citep{pap_lin84,bryden99,crida06,fung14}, 
while low mass planets may also open gaps if the disk viscosity is
small enough \citep{li09,dong11,duffell13}. Support for such disk-planet
interaction have begun to emerge in observations of circumstellar
disks that reveal annular gaps
\citep[e.g.][]{quanz13a,debes13,osorio14}, with possible evidence of
companions within them \citep[e.g.][]{quanz13b,reggiani14}.

%gaps can be unstable - vortices

A recent theoretical development in the study of planetary gaps is
their stability. When the disk viscosity is low and/or the planet mass
is large, the presence of potential vorticity (PV, the ratio of
vorticity to surface density) extrema can render planetary gaps
dynamically unstable due to what is now referred to as the `Rossby
wave instability' \citep[RWI,][]{lovelace99,li00}. This 
eventually leads to vortex formation 
\citep{li01,koller03,li05,valborro07}, which can significantly affect
orbital migration of the planet  \citep{ou07,li09,lin10}.

Vortex formation at gap edges may also have observable 
consequences. Because disc vortices represent pressure maxima, they are
able to collect dust particles 
\citep{barge95,inaba06,lyra13}. Dust-trapping at gap-edge vortices
have thus been suggested to explain asymmetric dust
distributions observed in several transition disks
\citep[e.g][]{marel13,isella13,perez14}.

However, studies of Rossby vortices at planetary gap-edges have 
adopted locally isothermal disks, %cite lin11, zhu14, lyra
where the disk temperature is a fixed function of
position only \citep[e.g.][]{lyra08,lin11a,zhu14,fu14}. 
%isothermal valid for outer disk  
On the other hand, the theory of the RWI was in fact
developed for adiabatic disks \citep{li00}, which permits
heating. %cite 
In adiabatic disks, the relevant quantity for stability
becomes a generalisation of the PV that include entropy variations
\citep{lovelace99}.

Gap-opening is associated with planet-induced spiral shocks. In an
isothermal disk, PV-generation across these isothermal shocks leads to
the RWI \citep{koller03,li05,valborro07,lin10}.    
However, if cooling is inefficient and the shock is non-isothermal,
then shock-heating may affect gap stability, since the
relevant quantity is an entropy-modified PV (described below), and
there is entropy-generation across the shocks.

For example, previous
linear simulations of the RWI found 
that increasing the sound-speed favours instability \citep{li00,lin13}. In the context of
planetary gaps, however, the increased temperature may also act to
stabilise the disk by making gap-opening more difficult. It is 
therefore of theoretical interest to clarify the effect of heating and
cooling on the stability of planetary gaps.

In this work, we extend the study of planetary gap stability against
vortex formation to non-isothermal discs. We include in the fluid energy
equation an one-parameter cooling prescription that allows us to probe
disk thermodynamics ranging from nearly isothermal to nearly
adiabatic.

This paper is organised as follows. In \S\ref{model} we describe the
equations governing the disc-planet system and initial conditions. Our
numerical approach, including diagnositic measures, are given in
\S\ref{method}. We present results from two sets of numerical
experiments. In \S\ref{linear1} we use disc-planet interaction to set
up discs with gaps, but study their stability without further
influence from the planet. %stability through background state 
We then examine the lifetime of gap-edge vortices in \S\ref{nonlinear}
as a function of the imposed cooling rate. We conclude and summarize
in \S\ref{summary} with a discussion of important caveats.

\input{model}
\input{results}
%\input{discussion}
\input{summary}

%\section*{Acknowledgments}

\bibliographystyle{mn2e}
\bibliography{ref}

\appendix
%\input{appendix}
%\input{appendix2}

\end{document}