### 2014-11-21 paper/summary.tex

\section{Summary and discussion}\label{summary}
In this paper, we have carried out numerical simulations of
non-isothermal disc-planet interaction.  
Our simulations were customised to examine the effect of a finite 
cooling time on the stability of gaps  
opened by giant planets to the so-called vortex or Rossby wave
instability. To do so, we  
included an energy equation with a cooling term that restores the 
disc temperature to its initial profile on a characteristic timescale
$t_c$. We studied the evolution of the gap stability as a function of 
$t_c$. This is a natural extension to previous studies of on gap
stability, which employ locally or strictly isothermal equations of
state.

We considered two types of numerical experiments. We first used
disc-planet interaction to self-consistently set up gap profiles,
which were then perturbed and evolved without further the influence of
the planet potential. This procedure isolates the effect of cooling on
gap stability through the set up of the initial gap profile. We find
that as the cooling time $t_c$ is increased, the gaps became more
stable, with lower growth rates of non-axisymmetric modes and the
dominant azimuthal wavenumber also decreases. This is consistent with
the notion that increasing $t_c$ leads to higher temperatures or $h$,
which opposes gap-opening by the planet. This means that the gaps
opened by the planet in a disc with longer $t_c$ is smoother and
therefore  more stable to the RWI.

In the second set of calculations, we included the planet potential
throughout the simulations  and examined the long-term evolution of
the gap-edge vortex that develops from the RWI. The vortex reaches 
a quasi-steady state lasting $O(10^3)$ orbits. Unlike the `planet-off'
simulations, in which vortices decay after linear growth and merging,
we find that with the planet potential kept on, the vortex amplitude
grows during this quasi-steady state, until it begins to induce
shocks, afterwhich the vortex amplitude begins to decay.

We observe the duration of the quasi-steady state increases with
increasing cooling time until some critical value, beyond which this
quasi-steady state shortens again. The timescale for the vortex to
decay, after reaching maximum amplitude, decreases monotonically with
increasing cooling time.

%technically we should probably say there is an optimum cooling time
%to maximise the quasi-steady state

Our results indicate there exists an optimum cooling timescale to
maximise the lifetime of gap-edge vortices. We suggest the
non-monotonic dependence of the quasi-steady state can be attributed
to the time required for the vortex to grow to sufficient amplitude to
induce shocks in the surrounding fluid, thereby losing energy and also
smooth out the gap edge.

For short cooling timescales, the planet is able to open a
deeper gap which favours the RWI, leading to stronger
vortices. However, for long  cooling timescales, the vortex grow
faster during the quasi-steady state {\bf check if statement true}. We
suggest the latter is due to the RWI being favoured with increasing
disc temperature, according to previous stability calculations. These
competing factors imply for both very short and long cooling
timescales, the vortex reaches its maximum amplitude, shock, and
begins to decay, sooner than intermediate cooling timescales.

%decay timescale depends on cooling time, but there seems to be a
%monotonic trend: longer cooling times, decay timescale is shorter, so
%this  
%doesn't contribute to non-monotonic trend

We remark that a non-monotonic dependence of the vortex lifetime was
also reported by \cite{fu14}, who performed isothermal disc-planet
simluations with different values of the   
disc aspect ratio. In their simulations the optimum aspect ratio is
$h=0.06$. In our simulations, the disc aspect ratio is a dynamical
variable, but by analysing the region where the vortex is located
($r-r_p\in[2,10]r_h$), % {\bf define this region}
we find for $\tilde\beta=5.0$ case, with the longest vortex lifetime, has
$h\approx0.06$ on average throughout its lifetime. 
{\bf if new definition of vortex lifetime gives other $\tilde{\beta}$ values as the maximum lifetime, then need to update this value of $h$}. 
Our result is therefore consistent with \citep{fu14}.

% we don't know how fu defined vortex lifetime

\subsection{Caveats and outlooks}
We discuss below several outstanding issues that needs to be addressed
in future works:

\emph{Convergence.} Although lower-resolution simulations performed in
the early stages of this project gave similar results  
(most importantly, the non-monotonic dependence of vortex lifetimes on
the cooling timescale), we did find the lower-resolution simulations  
typically yield longer vortex lifetimes than that reported in this
paper. This could be due to weaker RWI with low resolution.  
It will be neccessary to perform even higher resolution  simulations
in order to obtain quantitatively co nverged vortex lifetimes.

\emph{Self-gravity.} We have ignored disc self-gravity in this
study. Based on linear calculations, \cite{lovelace13} concluded
self-gravity to be important for the RWI when $Q<O(1/h)$, or
$Q\lesssim 20$ for $h\sim0.05$, as was typially considered in this
work. This suggests that self-gravity may affect vortex lifetimes even
when the Toomre $Q$ parameter is not small. In particular, given that
we observe vortices can reach significant over-densities (up to almost
an order of magnitude), it will be important to include disc
self-gravity in the future. %self-gravitational collapse of vortices

\emph{Three-dimensional (3D) effects.} A vortex in a 3D disc may be
subject to secondary instabilities that destroy them
\citep{lesur09,railton14}. This may be an important factor in
determining gap-edge vortex lifetimes in realistic discs. For example,
if these secondary instabilities sets in before the vortex grows to
sufficient amplitude to shock, then the dependence of the vortex
lifetime on the cooling timescale will be its effect through the 3D
instability (as opposed to the effect on the RWI itself, which is a 2D
instability). This problem needs to be clarified with full 3D
disc-planet simulations.

%3d is challenging because high resolution needed in for 3d instabilities
%However, the role of these instabilities on gap-edge vortices, where disc-planet  
%interaction can maintain the RWI, has not been clarified.