### 2015-02-15 paper/response.txt
revision in the works

- §§4 and 5: would the results of a simulation with a very short
  cooling time (beta << 1) be the same as those of a simulation with a
  locally isothermal equation of state and same unperturbed sound
  speed? 
  ********************************************************************
  We have performed addtional (low resolution) simulations with
  beta=0.01 to compare with a locally isothermal disks. We find a
  smaller linear growth rate in the locally isothermal case.

  ********************************************************************

- §4.4: why does the vortex’s perturbed density decrease on the long
  term? I suppose a related question is: what controls the long-term
  evolution of the vortex’s potential vorticity or entropy-modified
  potential vorticity? (whichever quantity is relevant). 
  ********************************************************************
  In theory a vortex should live forever in 2D inviscid flow. In
  practice, however, there is numerical viscosity which we suspect is
  responsible for the long term decay. As the vortex weakens
  (elongates) its area increases and its surface density drops.

  viscosity and elongation of vortex

  ask robert: does gap shape change significantly?

  ********************************************************************

- §5.2-5.3: here again I’d like to have more details on the vortex
        evolution and decay: 	
        . Do the authors see the vortex migrate? I did a simulation
        with the authors’ setup and a rather long cooling time
        (beta_tilde ~ 8) where I see the vortex moving from ~6.5 rh to
        ~5.3 rh relative to the planet’s orbital radius before the
        vortex starts decaying (rh = Hill radius). The time when the
        vortex starts decaying actually corresponds to a rather sudden
        increase in the planet-vortex distance, which goes up to ~6.8
        rh until the vortex dies off completely. If the authors get a
        similar migration behaviour in their simulations, maybe this
        could give further insight into the vortex evolution (vortex
        getting stronger during the quasi steady state before fading
        away). The connection between migration and the appearance of
        spiral waves launched by the vortex could also be made. 
	********************************************************************
        We thank the reviewer for raising the issue of vortex
        migration. We have re-examined our simulation results. We
        indeed observe vortex migration but our result is somewhat
        different to that of the reviewer.

 	We do not see significant vortex migration during quasi-steady
 	state. This may be due to the fact that the
 	vortex forms at a surface density maximum, which is able to
 	trap vortices (Paardekooper et al 2010).

	Instead, we observe the vortex begins to migrate inwards
	with the appearence of the spiral wakes (cf. planetary
	migration). However, we find the vortex eventually migrates
	outward to about 6.5rh before complete decay.

	We remark that for a vortex formed at a surface density
 	maximum (e.g. from the RWI), that subsequent perturbations may
 	lead to inwards or outwards vortex migration, depending on
 	what side of the maximum the vortex is displaced to
 	(Paardekooper et al 2010).

	Although we do not understand the discrepency between our
	results and the reviewer's, we do agree with the reviewer
	that vortex migration probably plays a role in determining
	the overall vortex lifetime (it may affect the duration of the
	quasi-steady state and/or the decay timescale). We now mention
	this in the discussions and caveats.

        ********************************************************************

        . During the quasi steady state the vortex grows as a result
        of continuous vortensity generation upon crossing the planet’s
        wake. What about possible turbulent diffusion of potential
        vorticity if the planet forms in a region of the disc with
        non-negligible ‘turbulent viscosity’? Have the authors done
        simulations with a non vanishingly small viscosity? 
        ********************************************************************
	Since a low-viscosity disc is needed for the RWI to operate,
        we have chosen to focus on invisicd simulations. This also
        helps to minimize vortex decay directly due to a physical
        viscosity, so that we can examine the influence of
        thermodynamics.

	Thus, we expect that vortex lifetimes in realistic viscous discs
	to be shorter than that seen in our 
	simulations. We mention this in the discussions.

        ********************************************************************
        . I was a bit surprised that the RWI couldn’t be launched
        again after the vortex had decayed, and I’d like the authors
        to be more specific about why it is so. About the vortex decay
        providing an effective alpha viscosity~ 1d-2: isn’t this a
        time-varying quantity, decreasing as the vortex is dying? If
        so is it really the reason why the RWI isn’t launched again?
        What about the disc’s aspect ratio at the outer gap edge after
        the vortex’s full decay?     
        ********************************************************************
	The quoted value of 0.01 is a typical viscosity associated
        with the vortex, averaged over radius and time.

	By the time the vortex dies out that there is
	no longer a `vortex viscosity', but the outer gap edge
	(surface density bump) has already been smoothed out, and the
	site is no longer suitable for the RWI. Note that from
	Fig. 11, smoothing out the surface density bump puts material
	into the outer disc.

	In order to re-launch the RWI, the surface density bump needs
	to reform. However, this seems difficult as the planet already
	resides in a deep gap with little material to clear out
	(c.f. the original surface density bump resulted from
	gap-opening: the planet pushes material to either side of its
	orbit).

        ********************************************************************

- §6.1: two caveats / outlooks could be added: (i) having a more
  realistic cooling model (same caveat as for numerical simulations of
  disc fragmentation by the gravitational instability…) and (ii)
  considering planet migration. 
  ********************************************************************
  These are now added. 
  ********************************************************************
- §2: given the very low shear viscosity used in this work, I would
  emphasize straight after eqs. (2) and (3) that f_v and H include the
  contribution of a (Von Neumann-Ritchmyer) artificial bulk viscosity
  to handle shocks. 
  ********************************************************************
  These comments are now added. 
  ********************************************************************
- §4.2: the minimum value of 2d-4 for kappa^2 + N^2 should rather be
  expressed in units of Omega^2 at the planet’s orbital radius. Same
  advice for N, for which the value of 1d-5 didn’t really speak to me
  until I realised that the planet’s orbital radius wasn’t the code’s
  unit of length. By the way, where is N evaluated? 
  ********************************************************************
  The frequencies are now expressed in units of the orbital frequency
  at the planet's orbital radius. 
  ********************************************************************
- Fig. 6: the colour scale suggests that there is a discontinuous jump
  in perturbed density at orbital radii R > 40 r_in, where the density
  seems to take very small values… why is that? 
  ********************************************************************
  This is due to the open boundary conditions adopted at the outer
  boundary, which leads to mass loss. We now mention this in the
  caption of Fig. 6. 
  ********************************************************************